Follow-up Samples

### 2016-12-06 submission/manuscript_outline_20161024.Rmd

***Outcome of Model Training***
The range of the AUC for model training ranged from a minimum of ```r format(model_test_AUC_summary["worse", "ROC_cv"], digits = 3)``` to a maximum of ```r format(model_test_AUC_summary["best", "ROC_cv"], digits = 3)```.  To be conservative the model chosen for prediction of disease in the follow up samples had an AUC in the middle of all the ```r length(rownames(model_train_summary))``` runs which was ```r format(model_test_AUC_summary["middle", "ROC_cv"], 
digits = 3)```.  Interestingly, the worst AUC model from training performed the best on it's respective 80/20 split test data [Figure 3].  In fact the 80/20 test performance showed that the AUC for the middle model chosen was the most stable (best training model test set AUC = ```r format(model_test_AUC_summary["best", "AUC"], digits = 3)```, middle training model test set AUC = ```r format(model_test_AUC_summary["middle", "AUC"], digits = 3)```, worse training model test set AUC = ```r format(model_test_AUC_summary["worse", "AUC"], digits = 3)```).  That is to say it had the smallest change in AUC in comparison to the other minimum and maximum AUC trained models.  There was no significant difference between the AUC of the best and middle training models (P-value = ```r format(roc_pvalues["BH_adj", "best_vs_middle"], digits = 3)```).  There was a significant difference between the worse training model and both the best training model (P-value = ```r format(roc_pvalues["BH_adj", "best_vs_worse"], digits = 3)```) and middle training model (P-value = ```r format(roc_pvalues["BH_adj", "middle_vs_worse"], digits = 3)```).

@@

Naturally, it is curious that normal staples of many screening studies such as Fusobacterium, Parvimonas, and Peptostreptococcus were not present in the majority of the training models.  One potential explanation for this is that FIT provides the same information to the model as these three organisms and so the model uses FIT preferentially over them.  This has been suggested to be the case in a previous study (**insert Baxter Study**).  Regardless, our study does not argue against the importance of these bacterium in CRC initiation or pathogenesis but rather that the best model does not utilize these specific bacteria for prediction purposes.

>>>>

***Outcome of Model Training***
The range of the AUC for model training ranged from a minimum of ```r format(model_test_AUC_summary["worse", "ROC_cv"], digits = 3)``` to a maximum of ```r format(model_test_AUC_summary["best", "ROC_cv"], digits = 3)```.  To be conservative the model chosen for prediction of disease in the follow up samples had an AUC in the middle of all the ```r length(rownames(model_train_summary))``` runs which was ```r format(model_test_AUC_summary["middle", "ROC_cv"], 
digits = 3)```.  Interestingly, the worst AUC model from training performed the best on it's respective 80/20 split test data [Figure 3].  In fact the 80/20 test performance showed that the AUC for the middle model chosen was the most stable (best training model test set AUC = ```r format(model_test_AUC_summary["best", "AUC"], digits = 3)```, middle training model test set AUC = ```r format(model_test_AUC_summary["middle", "AUC"], digits = 3)```, worse training model test set AUC = ```r format(model_test_AUC_summary["worse", "AUC"], digits = 3)```).  That is to say it had the smallest change in AUC in comparison to the other minimum and maximum AUC trained models.  There was no significant difference between the AUC of the best and middle training models (P-value = ```r format(roc_pvalues["best", "middle"], digits = 3)```).  There was also no difference in the middle model versus worse (P-value = ```r format(roc_pvalues["middle", "worse"], digits = 3)```) or full data model (P-value = ```r format(roc_pvalues["middle", "full"], digits = 3)```).  The two comparisons with a significant difference were between the worse training model and the best training model (P-value = ```r format(roc_pvalues["best", "worse"], scientific = TRUE, digits = 3)```) and the worse training model and full data model (P-value = ```r format(roc_pvalues["worse", "full"], scientific = TRUE, digits = 3)```).

@@

Naturally, it is curious that normal staples of many screening studies such as Fusobacterium, Parvimonas, and Peptostreptococcus were not present in the majority of the training models.  One potential explanation for this is that FIT provides the same information to the model as these three organisms and so the model uses FIT preferentially over them.  This has been suggested to be the case in a previous study (**insert Baxter Study**).  Regardless, our study does not argue against the importance of these bacterium in CRC initiation or pathogenesis but rather that the best model does not utilize these specific bacteria for prediction purposes.  Another potential reason why we did not identify the usual suspects is that they may not change much between our initial and follow up samples.  That is to say that they are consistently present even after removal of the lesion but surgery.

### 2016-12-06 submission/manuscript_outline_20161024.tex

\textbf{\emph{Outcome of Model Training}} The range of the AUC for model
training ranged from a minimum of \texttt{0.795} to a maximum of
\texttt{0.854}. To be conservative the model chosen for prediction of
disease in the follow up samples had an AUC in the middle of all the
\texttt{100} runs which was \texttt{0.819}. Interestingly, the worst AUC
model from training performed the best on it's respective 80/20 split
test data {[}Figure 3{]}. In fact the 80/20 test performance showed that
the AUC for the middle model chosen was the most stable (best training
model test set AUC = \texttt{0.682}, middle training model test set AUC
= \texttt{0.815}, worse training model test set AUC = \texttt{0.931}).
That is to say it had the smallest change in AUC in comparison to the
other minimum and maximum AUC trained models. There was no significant
difference between the AUC of the best and middle training models
(P-value = \texttt{0.0698}). There was a significant difference between
the worse training model and both the best training model (P-value =
\texttt{0.000397}) and middle training model (P-value =
\texttt{0.0444}).

\textbf{\emph{Most Important Variables to the Model}} Overall, there
were a total of \texttt{32} variables identified as being present in
more than 50\% of the training models {[}Table S2{]}. The top 5 most
important bacterial OTUs were \texttt{Lachnospiraceae\ (Otu000013)},
\texttt{Escherichia/Shigella\ (Otu000018)},
\texttt{Ruminococcaceae\ (Otu000020)},
\texttt{Ruminococcus\ (Otu000017)}, and
\texttt{Porphyromonas\ (Otu000153)}. These 5 OTUs were present in at
least \texttt{99} out of the total \texttt{100} different 80/20 runs.

>>>>

\textbf{\emph{Outcome of Model Training}} The range of the AUC for model
training ranged from a minimum of \texttt{0.795} to a maximum of
\texttt{0.854}. To be conservative the model chosen for prediction of
disease in the follow up samples had an AUC in the middle of all the
\texttt{100} runs which was \texttt{0.819}. Interestingly, the worst AUC
model from training performed the best on it's respective 80/20 split
test data {[}Figure 3{]}. In fact the 80/20 test performance showed that
the AUC for the middle model chosen was the most stable (best training
model test set AUC = \texttt{0.682}, middle training model test set AUC
= \texttt{0.815}, worse training model test set AUC = \texttt{0.931}).
That is to say it had the smallest change in AUC in comparison to the
other minimum and maximum AUC trained models. There was no significant
difference between the AUC of the best and middle training models
(P-value = \texttt{0.419}). There was also no difference in the middle
model versus worse (P-value = \texttt{0.178}) or full data model
(P-value = \texttt{1}). The two comparisons with a significant
difference were between the worse training model and the best training
model (P-value = \texttt{7.94e-04}) and the worse training model and
full data model (P-value = \texttt{3.52e-03}).

\textbf{\emph{Most Important Variables to the Model}} Overall, there
were a total of \texttt{32} variables identified as being present in
more than 50\% of the training models {[}Table S2{]}. The top 5 most
important bacterial OTUs were \texttt{Bacteria\ (Otu000013)},
\texttt{Escherichia/Shigella\ (Otu000018)},
\texttt{Bacteria\ (Otu000020)}, \texttt{Ruminococcus\ (Otu000017)}, and
\texttt{Porphyromonas\ (Otu000153)}. These 5 OTUs were present in at
least \texttt{99} out of the total \texttt{100} different 80/20 runs.
